Efficient Handling of String-Number Conversion

### 2019-11-23 draft.tex
sec 5

\begin{abstract}
String-number conversion is an important class of constraints needed for the symbolic execution of string-manipulating programs. In particular solving string constraints with string-number conversion  is necessary  for the analysis of scripting languages such as JavaScript and Python, where string-number conversion is a part of the definition of the core semantics of these languages. However, the state-of-the-art constraint solvers have only very limited support to this type of constraint. We propose to overcome these limitations with an approach that can efficiently support both string-number conversion and other common types of string constraints. Experimental results show that it significantly outperforms other state-of-the-art tools on benchmarks that involves string-number conversion.
\end{abstract}
 \setcopyright{none} 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newcommand{\hide}[1]{}
\newcommand{\tool}{{\textsf{Z3-PFA}}}
\newcommand{\dr}{\mathcal{R}}
\newcommand{\nat}{\mathbb{N}}
\newcommand{\integers}{\mathbb{Z}}
\newcommand{\todo}[1]{{\color{blue}TODO: #1}}
\newcommand{\lh}[1]{{\color{orange}Lukas: #1}}
\newcommand{\changed}[1]{{\color{orange}#1}}
\newcommand{\yfc}[1]{{\color{blue}YFC: #1}}
\newcommand{\petr}[1]{{\color{pink}Petr: #1}}
\newcommand{\chatAt}[2]{\mbox{\textsf{charAt}($#1$, $#2$)}}
\newcommand{\ite}[3]{\mbox{\textsf{ite}($#1$, $#2$, $#3$)}}
\newcommand{\sti}[1]{\mbox{\textsf{toNum}($#1$)}}
\newcommand{\its}[1]{\mbox{\textsf{toStr}($#1$)}}
\newcommand{\varn}{\mbox{$V$}}
\newcommand{\vars}{\mbox{$X$}}
\newcommand{\cvars}{\mbox{$\mathbb{V}_{\Sigma_\epsilon}$}}
\newcommand{\pvars}{\mbox{$\mathbb{V}_{\sharp}$}}
%\newcommand{\cvar}{\mbox{$V_{\Sigma_\epsilon}$}}
%\newcommand{\cvarone}{\mbox{$V_{\Sigma_\epsilon}$}}
%\newcommand{\cvartwo}{\mbox{$V'_{\Sigma_\epsilon}$}}
\newcommand{\cvarone}{V}
\newcommand{\cvartwo}{V'}
\newcommand{\cvar}{V}
\newcommand{\modelsof}[1]{[\![#1]\!]}
\newcommand{\true}{\mbox{$\mathsf{true}$}}
\newcommand{\false}{\mbox{$\mathsf{false}$}}
\newcommand{\enc}[1]{[\![#1]\!]}
%\newcommand{\parikhof}[1]{\mathit{Parikh}{(#1)}}
\newcommand{\parikhof}[1]{\mathbb{P}{(#1)}}
\newcommand{\parikhwof}[2]{|#1|_{#2}}
%\newcommand{\semof}[1]{||#1||}
\newcommand{\semof}[1]{\modelsof{#1}}
\newcommand{\decode}[1]{\mathit{decode_{#1}}}
%\newcommand{\parikhfof}[1]{\Phi_{\parikhof{#1}}}
\newcommand{\parikhfof}[1]{\Phi_{\mathbb{P}}(#1)} %shouldn't it be something that takes automaton as parameter and returns formula?
\newcommand{\pim}{I_{\#}}
%\newcommand{\syncop}{\Cap}
\newcommand{\syncop}{\times}
\newcommand{\syncof}[2]{#1 \syncop #2}
\newcommand{\syncfof}[2]{\Psi_{\syncof {#1} {#2}}}
\newcommand{\syncT}{T_\syncop}
\newcommand{\pvarsof}[1]{\#{#1}}
\newcommand{\pvar}{\pvarsof V}
\newcommand{\pvarone}{\pvarsof V}
\newcommand{\pvartwo}{\pvarsof {V'}}
\newcommand{\defeq}{::=}
\newcommand{\encode}[1]{\mathit{encode}_{#1}}
\newcommand{\eqwrt}[1]{=_{#1}}
\newcommand{\iequiv}{\equiv}
\newcommand{\restrict}[2]{#1_{#2}}
\newcommand{\underf}[2]{\mathit{under}_#1(#2)}
\newcommand{\paf}{\psi}

@@

We say that a mapping $I_e:\cvar\cup\#\cvar \rightarrow \nat$ is a \emph{word encoding} a word $w$  (or a $\pa$-encoding of $w$) if $w$ is an instantiation of some parametric word $x \in L(\pa)$ whose Parikh image and interpretation of character variables are defined by $I_e$.  
%
Conversely, $w$ is a \emph{$\pa$-decoding} of $I_e$.
%
We use $\encode \pa(w)$ below to denote all $\pa$-encodings of a word $w$,
and $\decode \pa (I_e)$ to denote all $\pa$-decodings of a word encoding $I_e$. Namely,  
% to \emph{word encodings} $e:\cvar\cup\#\cvar \rightarrow \nat$ via the PA $\pa$ and the function $\decode \pa$ does it in the reverse direction.
	$$\encode \pa( w ) \defeq \{I_e \mid  x \in L(\pa) , I(x) =w, I\in\semof \paf, I_e = I \cup \parikhof{x}\}$$
	$$\decode \pa(I_e) \defeq\{ w \mid x \in L(\pa) , I(x) =w, I\in\semof \paf, I_e = I \cup \parikhof{x} \}$$

@@

\changed{
The correctness of the entire construction of $\underf \dr {\insc}$ is expressed by  Theorem~\ref{theorem:correct}. 
It uses the decoding function $\decode \dr$ parameterized by the domain restriction $\dr$. 
The function takes as an argument an $\dr$-encoding of a string constraint, an interpretation $I_e \cup I_z$ consisting of a \emph{many-word encoding} $I_e: V_\dr\cup\#V_\dr\rightarrow \nat$ and an integer variable interpretation $I_z$. 
Intuitively, the many-word encoding $I_e$ contains for each $x\in X$ a $\dr(x)$-encoding of its value. 
The function then returns the set of string constraint interpretations 
$\decode  \dr (I_e\cup I_z) \defeq \{I_s \cup I_z \mid \forall x\in X: I_s(x) = \decode {\dr(x)} ((I_e)_{V_{\dr(x)}\cup \#V_{\dr(x)}})\}$.
The condition says that each string variable $x$ is interpreted as the ${\dr(x)}$-decoding of the word encoding $(I_e)_{V_{\dr(x)}\cup \#V_{\dr(x)}}$.
%We define the output of $\decode  \dr (I_e\cup I_z)$ as a set so that \emph{invalid word encodings}, i.e. those do not corresponds to possible word encodings from PFAs, are mapped to empty sets and valid ones are mapped to singleton sets.
We also define the $\dr$-encoding function as the counterpart of $\dr$-decoding,  
%
namely, for a model $I_s \cup I_z$ of a string constraint,  
we let $\encode \dr (I_s \cup I_z) =\{I_e\cup I_z\mid \decode \dr (I_e\cup I_z) = I_s \cup I_z \}$.
We lift $\decode \dr$ and $\encode \dr$ to sets of interpretations in the standard manner.
}

The correctness of the entire construction of $\underf \dr {\insc}$ is expressed by  Theorem~\ref{theorem:correct}. 
It uses the decoding function $\decode \dr$ parameterized by the domain restriction $\dr$. 
The $\decode \dr$ function maps a \emph{ many-words encoding} $I_e: V_\dr\cup\#V_\dr\rightarrow \nat$ paired with an interpretation of integer variables $I_z$ to a model $I_s \cup I_z$ of a string constraint. Note that the many-words encoding $I_e$ is a union of ${\dr(x)}$-encodings for all $x$ in $X$.
Formally, 
$\decode  \dr (I_e\cup I_z) = \{I_s \cup I_z \mid \forall x\in X: I_s(x) = \decode {\dr(x)} ((I_e)_{V_{\dr(x)}\cup \#V_{\dr(x)}})\}$.
The condition says each string variable $x$ is interpreted as the ${\dr(x)}$-decoding of the word encoding $(I_e)_{V_{\dr(x)}\cup \#V_{\dr(x)}}$.
%We define the output of $\decode  \dr (I_e\cup I_z)$ as a set so that \emph{invalid word encodings}, i.e. those do not corresponds to possible word encodings from PFAs, are mapped to empty sets and valid ones are mapped to singleton sets.
We define the $\dr$-encoding function as the counterpart of $\dr$-decoding,  
%
namely, for a model $I_s \cup I_z$ of a string constraint,  
we let $\encode \dr (I_s \cup I_z) =\{I_e\cup I_z\mid \decode \dr (I_e\cup I_z) = I_s \cup I_z \}$.
We lift $\decode \dr$ and $\encode \dr$ to sets of interpretations in the standard manner.
\begin{theorem}\label{theorem:correct}
$\decode \dr (\semof {\underf \dr {\insc}}) = \semof{\insc}^\dr$ 
\end{theorem}

>>>>

\begin{abstract}
String-number conversion is an important class of constraints needed for the symbolic execution of string-manipulating programs. In particular solving string constraints with string-number conversion  is necessary  for the analysis of scripting languages such as JavaScript and Python, where string-number conversion is a part of the definition of the core semantics of these languages. However, the state-of-the-art constraint solvers have only very limited support to this type of constraint. We propose to overcome these limitations with an approach that can efficiently support both string-number conversion and other common types of string constraints. Experimental results show that it significantly outperforms other state-of-the-art tools on benchmarks that involves string-number conversion.
\end{abstract}
 \setcopyright{none} 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newcommand{\hide}[1]{}
\newcommand{\tool}{{\textsf{Z3-PFA}}}
\newcommand{\dr}{\mathcal{R}}
\newcommand{\nat}{\mathbb{N}}
\newcommand{\integers}{\mathbb{Z}}
\newcommand{\todo}[1]{{\color{blue}TODO: #1}}
\newcommand{\lh}[1]{{\color{orange}Lukas: #1}}
\newcommand{\changed}[1]{{\color{orange}#1}}
\newcommand{\yfc}[1]{{\color{blue}YFC: #1}}
\newcommand{\petr}[1]{{\color{pink}Petr: #1}}
\newcommand{\chatAt}[2]{\mbox{\textsf{charAt}($#1$, $#2$)}}
\newcommand{\ite}[3]{\mbox{\textsf{ite}($#1$, $#2$, $#3$)}}
\newcommand{\sti}[1]{\mbox{\textsf{toNum}($#1$)}}
\newcommand{\its}[1]{\mbox{\textsf{toStr}($#1$)}}
\newcommand{\varn}{\mbox{$Z$}}
\newcommand{\vars}{\mbox{$X$}}
\newcommand{\cvars}{\mbox{$\mathbb{V}_{\Sigma_\epsilon}$}}
\newcommand{\pvars}{\mbox{$\mathbb{V}_{\sharp}$}}
%\newcommand{\cvar}{\mbox{$V_{\Sigma_\epsilon}$}}
%\newcommand{\cvarone}{\mbox{$V_{\Sigma_\epsilon}$}}
%\newcommand{\cvartwo}{\mbox{$V'_{\Sigma_\epsilon}$}}
\newcommand{\cvarone}{V}
\newcommand{\cvartwo}{V'}
\newcommand{\cvar}{V}
\newcommand{\modelsof}[1]{[\![#1]\!]}
\newcommand{\true}{\mbox{$\mathsf{true}$}}
\newcommand{\false}{\mbox{$\mathsf{false}$}}
\newcommand{\enc}[1]{[\![#1]\!]}
%\newcommand{\parikhof}[1]{\mathit{Parikh}{(#1)}}
\newcommand{\parikhof}[1]{\mathbb{P}{(#1)}}
\newcommand{\parikhwof}[2]{|#1|_{#2}}
%\newcommand{\semof}[1]{||#1||}
\newcommand{\semof}[1]{\modelsof{#1}}
\newcommand{\decode}[1]{\mathit{decode_{#1}}}
%\newcommand{\parikhfof}[1]{\Phi_{\parikhof{#1}}}
\newcommand{\parikhfof}[1]{\Phi_{\mathbb{P}}(#1)} %shouldn't it be something that takes automaton as parameter and returns formula?
\newcommand{\pim}{I_{\#}}
%\newcommand{\syncop}{\Cap}
\newcommand{\syncop}{\times}
\newcommand{\syncof}[2]{#1 \syncop #2}
\newcommand{\syncfof}[2]{\Psi_{\syncof {#1} {#2}}}
\newcommand{\syncT}{T_\syncop}
\newcommand{\pvarsof}[1]{\#{#1}}
\newcommand{\pvar}{\pvarsof V}
\newcommand{\pvarone}{\pvarsof V}
\newcommand{\pvartwo}{\pvarsof {V'}}
\newcommand{\defeq}{::=}
\newcommand{\encode}[1]{\mathit{encode}_{#1}}
\newcommand{\eqwrt}[1]{=_{#1}}
\newcommand{\iequiv}{\equiv}
\newcommand{\restrict}[2]{#1_{#2}}
\newcommand{\underf}[2]{\mathit{under}_#1(#2)}
\newcommand{\paf}{\psi}

@@

We say that a mapping $I_e:\cvar\cup\#\cvar \rightarrow \nat$ is a \emph{word encoding} of a word $w$  (or a $\pa$-encoding of $w$) if $w$ is an instantiation of some parametric word $x \in L(\pa)$ whose Parikh image and interpretation of character variables are defined by $I_e$.  
%
Conversely, $w$ is a \emph{$\pa$-decoding} of $I_e$.
%
We use $\encode \pa(w)$ below to denote all $\pa$-encodings of a word $w$,
and $\decode \pa (I_e)$ to denote all $\pa$-decodings of a word encoding $I_e$. Namely,  
% to \emph{word encodings} $e:\cvar\cup\#\cvar \rightarrow \nat$ via the PA $\pa$ and the function $\decode \pa$ does it in the reverse direction.
	$$\encode \pa( w ) \defeq \{I_e \mid  x \in L(\pa) , I(x) =w, I\in\semof \paf, I_e = I \cup \parikhof{x}\}$$
	$$\decode \pa(I_e) \defeq\{ w \mid x \in L(\pa) , I(x) =w, I\in\semof \paf, I_e = I \cup \parikhof{x} \}$$

@@

The correctness of the entire construction of $\underf \dr {\insc}$ is expressed by  Theorem~\ref{theorem:correct}. 
It uses the decoding function $\decode \dr$ parameterized by the domain restriction $\dr$. 
Let $Z$ be the set of integer variables in $\insc$.
The function maps an interpretations $I_e$ over $Z\cup V_\dr\cup\#V_\dr$
to an interpretation over $Z \cup X$, if exists. Informally, it ``decodes" an interpretation of integer variables $Z\cup V_\dr\cup\#V_\dr$ to an interpretation of variables in the string constraint $\insc$. Formally,
we define $\decode  \dr (I_e) \defeq \{I \mid \forall z\in Z: I(z) = I_e(z) \wedge\forall x\in X: \{I(x)\} = \decode {\dr(x)} ((I_e)_{V_{\dr(x)}\cup \#V_{\dr(x)}})\}$.
The condition says that (1) $I$ and $I_e$ are consistent over the $Z$ and (2) $(I_e)_{V_{\dr(x)}\cup \#V_{\dr(x)}}$ is a word encoding that $\dr(x)$-encodes $I(x)$.
%We define the output of $\decode  \dr (I_e\cup I_z)$ as a set so that \emph{invalid word encodings}, i.e. those do not corresponds to possible word encodings from PFAs, are mapped to empty sets and valid ones are mapped to singleton sets.
We also define the $\dr$-encoding function as the counterpart of $\dr$-decoding,  
%
namely, for a interpretation $I$ of the string constraint $\insc$,  
we let $\encode \dr (I) =\{I_e\mid \decode \dr (I_e) = I \}$.
We lift $\decode \dr$ and $\encode \dr$ to sets of interpretations in the standard manner.
\begin{theorem}\label{theorem:correct}
$\decode \dr (\semof {\underf \dr {\insc}}) = \semof{\insc}^\dr$ 
\end{theorem}
